Supplemental Material for: The evolution of gene expression in the term placenta of viviparous mammals

### 2016-09-01 manuscript/supplemental_material.Rnw
add short captions

\begin{figure}[hb]
<<transcriptome.mapping.figure,fig.width=8,fig.height=8,echo=FALSE>>=
star.logs.unmapped <- 
    star.logs[grepl("%",field) & 
                  grepl("unmapped|too many loci",field) & 
                      !grepl("too many mismatches",field),]
star.logs.unmapped <- 
    star.logs.unmapped[,field:=capfirst(gsub("\\% of reads (unmapped:|mapped to) ","",field))]
pushViewport(viewport(layout=grid.layout(ncol=1,nrow=2,heights=c(2,1))))
print(ggplot(star.logs.unmapped,
             aes(x=capfirst(species),y=as.numeric(gsub("%","",value))))
      + geom_boxplot()
      + facet_grid(field~.,scales="free")
      + ylab("Percentage of Reads Unmapped")
      + xlab(NULL)
      + theme_bw()
      + theme(legend.position="top",
              plot.margin=unit(c(0,0,0,1.1),"strwidth",data=list("","","","e8")),
              axis.title.x=element_blank(),
              axis.text.x = element_blank()),#element_text(angle = 30, hjust = 1, vjust=1,face="italic")),
      vp=viewport(layout.pos.col=1,layout.pos.row=1)
      )
pushViewport(viewport(layout.pos.col=1,layout.pos.row=1))
grid.text("A",
          y=unit(1,"npc")-unit(0.1,"cm"),
          x=unit(0,"npc")+unit(0.1,"cm"),
          gp=gpar(fontsize=24),
          just=c("left","top"))
popViewport()          
print(ggplot(star.logs[grepl("input reads$",field),],
             aes(x=capfirst(species),y=as.numeric(value)))
      + geom_boxplot()
      + scale_y_log10()
      + ylab("Reads")
      + xlab("Species")
      + theme_bw()
      + theme(legend.position="top",
              plot.margin=unit(c(0,1.6,0,0),"strheight",data=list("","TO","","")),
              axis.text.x = element_text(angle = 30, hjust = 1, vjust=1,face="italic")),
      vp=viewport(layout.pos.col=1,layout.pos.row=2))
pushViewport(viewport(layout.pos.col=1,layout.pos.row=2))
grid.text("B",
          y=unit(1,"npc")-unit(0.1,"cm"),
          x=unit(0,"npc")+unit(0.1,"cm"),
          gp=gpar(fontsize=24),
          just=c("left","top"))
popViewport(2)
@ 
\caption{RNAseq read and alignment statistics for all species. A)
  Boxplot of percentages of unmapped reads in each species; 0\% of
  reads in all species were unmapped because of mismatches. B) Boxplot
  of number of RNAseq reads in each species.}
\label{fig:transcriptome_statistics}
\end{figure}

@@

print(xtable(go.analysis.wide.table,
             display=c("d","s","s","s","d","e"),
             align=  c("c","c","c","c","c","c"),
             caption="GO terms with significant enrichment ($\\text{FDR}\\le0.05$)
  in 12 or more of the species analyzed from genes in the top 1\\% of
  expression. $n(\\text{FDR}\\le0.05)$ is the number of species with
  significant enrichment ($\\text{FDR}\\le 0.05$) in the GO term
  listed. $max(\\text{FDR}\\le0.05)$ is the maximum FDR value in all
  species with an FDR of less than $0.05$.",
  label="tab:go_terms_all_species"
  ),
  math.style.exponents=TRUE,
  include.rownames=FALSE,
  booktabs=TRUE,
  floating=FALSE,
  tabular.environment="longtable",
  caption.placement="top",
  size="\\footnotesize"
  )

@@

<<go.analysis.top.5p.all.species,echo=FALSE,results="asis">>=
for (this.species in names(species.names)) {
    cat(paste0("\\subsubsection{Significantly enriched ($\\text{FDR}\\le0.05$) GO terms in ",
                           " the top 1\\% of \\textit{",
                           capfirst(this.species),"} expressed transcripts}\n\n"))
    this.caption <- paste0("Significantly enriched ($\\text{FDR}\\le0.05$) GO terms in ",
                           " the top 1\\% of \\textit{",
                           capfirst(this.species),"} expressed transcripts.")
    print(xtable(go.results[grepl("Top 1% No Housekeeping",type) & FDR <= 0.05 & species == this.species,
                            list(GO.ID,Term,FDR,Ontology)
                            ][order(FDR)],
                 caption=this.caption,
                 label=paste0("tab:go_terms_enriched_top_one_",gsub(" ","_",this.species)),
                 display=c("s","s","s","e","s")
                 ),
          size="\\small",
          booktabs=TRUE, 
          math.style.exponents=TRUE,
          include.rownames=FALSE,
          tabular.environment="longtable",
          caption.placement="top",
          floating=FALSE)
}
@

@@

<<go.analysis.all.species,echo=FALSE,results="asis">>=
for (this.species in names(species.names)) {
    cat(paste0("\\subsubsection{Significantly enriched ($\\text{FDR}\\le0.05$) GO terms in ",
                           " the top 100 \\textit{",
                           capfirst(this.species),"} expressed transcripts}\n\n"))
    this.caption <- paste0("Significantly enriched ($\\text{FDR}\\le0.05$) GO terms in ",
                           " the top 100 \\textit{",capfirst(this.species),"} expressed transcripts")
    this.label <- paste0("tab:go_terms_enriched_",gsub(" ","_",this.species))
    print(xtable(go.results[grepl("100 No Housekeeping",type) & FDR <= 0.05 & species == this.species,
                            list(GO.ID,Term,FDR,Ontology)
                            ][order(FDR)],
                 caption=this.caption,
                 label=this.label,
                 display=c("s","s","s","e","s")
                 ),
          size="\\small",
          booktabs=TRUE,
          math.style.exponents=TRUE,
          include.rownames=FALSE,
          tabular.environment="longtable",
          caption.placement="top",
          floating=FALSE)
}
@

>>>>

\begin{figure}[hb]
<<transcriptome.mapping.figure,fig.width=8,fig.height=8,echo=FALSE>>=
star.logs.unmapped <- 
    star.logs[grepl("%",field) & 
                  grepl("unmapped|too many loci",field) & 
                      !grepl("too many mismatches",field),]
star.logs.unmapped <- 
    star.logs.unmapped[,field:=capfirst(gsub("\\% of reads (unmapped:|mapped to) ","",field))]
pushViewport(viewport(layout=grid.layout(ncol=1,nrow=2,heights=c(2,1))))
print(ggplot(star.logs.unmapped,
             aes(x=capfirst(species),y=as.numeric(gsub("%","",value))))
      + geom_boxplot()
      + facet_grid(field~.,scales="free")
      + ylab("Percentage of Reads Unmapped")
      + xlab(NULL)
      + theme_bw()
      + theme(legend.position="top",
              plot.margin=unit(c(0,0,0,1.1),"strwidth",data=list("","","","e8")),
              axis.title.x=element_blank(),
              axis.text.x = element_blank()),#element_text(angle = 30, hjust = 1, vjust=1,face="italic")),
      vp=viewport(layout.pos.col=1,layout.pos.row=1)
      )
pushViewport(viewport(layout.pos.col=1,layout.pos.row=1))
grid.text("A",
          y=unit(1,"npc")-unit(0.1,"cm"),
          x=unit(0,"npc")+unit(0.1,"cm"),
          gp=gpar(fontsize=24),
          just=c("left","top"))
popViewport()          
print(ggplot(star.logs[grepl("input reads$",field),],
             aes(x=capfirst(species),y=as.numeric(value)))
      + geom_boxplot()
      + scale_y_log10()
      + ylab("Reads")
      + xlab("Species")
      + theme_bw()
      + theme(legend.position="top",
              plot.margin=unit(c(0,1.6,0,0),"strheight",data=list("","TO","","")),
              axis.text.x = element_text(angle = 30, hjust = 1, vjust=1,face="italic")),
      vp=viewport(layout.pos.col=1,layout.pos.row=2))
pushViewport(viewport(layout.pos.col=1,layout.pos.row=2))
grid.text("B",
          y=unit(1,"npc")-unit(0.1,"cm"),
          x=unit(0,"npc")+unit(0.1,"cm"),
          gp=gpar(fontsize=24),
          just=c("left","top"))
popViewport(2)
@ 
\caption[RNAseq statistics]{RNAseq read and alignment statistics for all species. A)
  Boxplot of percentages of unmapped reads in each species; 0\% of
  reads in all species were unmapped because of mismatches. B) Boxplot
  of number of RNAseq reads in each species.}
\label{fig:transcriptome_statistics}
\end{figure}

@@

print(xtable(go.analysis.wide.table,
             display=c("d","s","s","s","d","e"),
             align=  c("c","c","c","c","c","c"),
             caption=c("GO terms with significant enrichment ($\\text{FDR}\\le0.05$)
  in 12 or more of the species analyzed from genes in the top 1\\% of
  expression. $n(\\text{FDR}\\le0.05)$ is the number of species with
  significant enrichment ($\\text{FDR}\\le 0.05$) in the GO term
  listed. $max(\\text{FDR}\\le0.05)$ is the maximum FDR value in all
  species with an FDR of less than $0.05$.","GO terms in 12 or more species"),
  label="tab:go_terms_all_species"
  ),
  math.style.exponents=TRUE,
  include.rownames=FALSE,
  booktabs=TRUE,
  floating=FALSE,
  tabular.environment="longtable",
  caption.placement="top",
  size="\\footnotesize"
  )

@@

<<go.analysis.top.5p.all.species,echo=FALSE,results="asis">>=
for (this.species in names(species.names)) {
    cat(paste0("\\subsubsection{Significantly enriched ($\\text{FDR}\\le0.05$) GO terms in ",
                           " the top 1\\% of \\textit{",
                           capfirst(this.species),"} expressed transcripts}\n\n"))
    this.caption <- paste0("Significantly enriched ($\\text{FDR}\\le0.05$) GO terms in ",
                           " the top 1\\% of \\textit{",
                           capfirst(this.species),"} expressed transcripts.")
    print(xtable(go.results[grepl("Top 1% No Housekeeping",type) & FDR <= 0.05 & species == this.species,
                            list(GO.ID,Term,FDR,Ontology)
                            ][order(FDR)],
                 caption=c(this.caption,paste0("GO terms from \\textit{",
                                               capfirst(this.species),"} top 1\\%")),
                 label=paste0("tab:go_terms_enriched_top_one_",gsub(" ","_",this.species)),
                 display=c("s","s","s","e","s")
                 ),
          size="\\small",
          booktabs=TRUE, 
          math.style.exponents=TRUE,
          include.rownames=FALSE,
          tabular.environment="longtable",
          caption.placement="top",
          floating=FALSE)
}
@

@@

<<go.analysis.all.species,echo=FALSE,results="asis">>=
for (this.species in names(species.names)) {
    cat(paste0("\\subsubsection{Significantly enriched ($\\text{FDR}\\le0.05$) GO terms in ",
                           " the top 100 \\textit{",
                           capfirst(this.species),"} expressed transcripts}\n\n"))
    this.caption <- paste0("Significantly enriched ($\\text{FDR}\\le0.05$) GO terms in ",
                           " the top 100 \\textit{",capfirst(this.species),"} expressed transcripts")
    this.label <- paste0("tab:go_terms_enriched_",gsub(" ","_",this.species))
    print(xtable(go.results[grepl("100 No Housekeeping",type) & FDR <= 0.05 & species == this.species,
                            list(GO.ID,Term,FDR,Ontology)
                            ][order(FDR)],
                 caption=c(this.caption,paste0("GO terms from \\textit{",
                                               capfirst(this.species),"} top 100")),
                 label=this.label,
                 display=c("s","s","s","e","s")
                 ),
          size="\\small",
          booktabs=TRUE,
          math.style.exponents=TRUE,
          include.rownames=FALSE,
          tabular.environment="longtable",
          caption.placement="top",
          floating=FALSE)
}
@
